---
title: "Crime"
author: "datadavidz"
date: "January 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of Motor Vehicle Theft in Chicago

A subset of criminal data recorded by the FBI is analyzed in this report.  The data is for motor vehicle thefts in Chicago from 2001 to 2012. A description of the variables follows:
  
* **ID:** a unique identifier for each observation
* **Date:** the date the crime occurred
* **LocationDescription:** the location where the crime occurred
* **Arrest:** whether or not an arrest was made for the crime (TRUE if an arrest was made and FALSE if no arrest made)
* **Domestic:** whether or not the crime was a domestic crime meaning that it was committed against a family member (TRUE if it was domestic and FALSE if it was not domestic)
* **Beat:** the area or "beat" in which the crime occurred. This is the smallest regional division defined by the Chicago police department
* **District:** the police district in which the crime occurred.  Each district is composed of many beats and are defined by the Chicago Police Department.
* **CommunityArea:** the community area in which the crime occurred.  Since the 1920s, Chicago has been divided into what are called "community areas" of which there are now 77.  The community areas were devised in an attempt to create socially homogeneous regions.
* **Year:** the year in which the crime occurred
* **Latitude:** the latitude of the location at which the crime occurred
* **Longitude:** the longitude of the location at which the crime occurred

###Loading the Data
Read the dataset mvtWeek1.csv into a dataframe called "mvt".

```{r load}
setwd("~/Crime")
mvt <- read.csv("mvtWeek1.csv")
```

###Display str information

```{r str}
str(mvt)
```

###Exploring the data variables

The maximum ID number is `r max(mvt$ID)`.
```{r explore}
max(mvt$ID)
min(mvt$Beat)
sum(mvt$Arrest == TRUE)
sum(mvt$LocationDescription == "ALLEY")
```

###Make the date variable an R date

The dates are in Month/Day/Year Hour:Minute format.  For example, the first date entry is `r mvt$Date[1]`.

```{r}
DateConvert = as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))
median(DateConvert)

#Extract Month and Day and add to mvt dataframe
mvt$Month <- months(DateConvert)
mvt$Day <- weekdays(DateConvert)
#Replace original date variable with converted dates
mvt$Date <- DateConvert
```
Motor vehicle crimes by month:
```{r}
table(mvt$Month)
```
The month with the fewest motor vehicle thefts was February.

Motor vehicle thefts by weekday:
```{r}
table(mvt$Day)
```
The weekday with the most motor vehicle thefts was Friday.

#Visualizing the dataset
```{r}
hist(mvt$Date, breaks=100)
```

Motor vehicle thefts appear to be trending down over this time period.

```{r}
boxplot(mvt$Date ~ mvt$Arrest)
```

More arrests appear to be occurring in the beginning of the time period.

Table of Crimes with and without Arrest by Year
```{r}
table(mvt$Year, mvt$Arrest)
```

The proportion of crimes with arrest made in 2001 was `r 2152/(18517+2152)`.  The proportion of crimes with arrest made in 2007 was `r 1212/(13068+1212)`.

Table of Motor Vehicle Thefts by Location
```{r}
sort(table(mvt$LocationDescription))
```

The top 5 locations are: STREET, PARKING LOT/GARAGE(NON.RESID.), ALLEY, GAS STATION and DRIVEWAY - RESIDENTIAL.
```{r}
topLocations <- c("STREET", "PARKING LOT/GARAGE(NON.RESID.)", "ALLEY", "GAS STATION", "DRIVEWAY - RESIDENTIAL")
top5 <- subset(mvt, LocationDescription %in% topLocations)

# Remove the unused factor variables from LocationDescription
top5$LocationDescription <- factor(top5$LocationDescription)
```

Table of Arrests for Top 5 Locations
```{r}
table(top5$LocationDescription, top5$Arrest)
```
Arrest rate at GAS STATION is highest `r 439/(439+1672)`.

Table of Motor Vehicle Thefts by Weekday
```{r}
table(top5$LocationDescription, top5$Day)
```

The most motor vehicle thefts at GAS STATION happen on Saturday.
The fewest motor vehicle thefts at DRIVEWAY - RESIDENTIAL happen on Saturday.







